---
title: "EDA Assignment Prep"
author: "Amber Camp"
format: html
editor: visual
---

## EDA Assignment Prep: Pokemon data

### Libraries

```{r}
library(here)
library(tidyverse)
library(janitor)
```

### Data

```{r}
pokemon <- read_csv(here("data/pokemon.csv"))

pokemon <- pokemon %>% 
  clean_names()
```

### Describe the data: Pokemon 

Data author: Alberto Barradas

Description: This is pokemon data for 721 pokemon and their stats.

How many observations (rows) are there? 800

How many variables (columns) are there? 13

### skimr: Quick overview

```{r}
skimdata <- skimr::skim(pokemon)
```

### Summary: Quick overview

```{r}
summary(pokemon)
```

### Categorical vs. Continuous

```{r}

pokemon$number <- as.factor(pokemon$number)
#class(pokemon$number)
#summary(pokemon$number)

pokemon$type_1 <- as.factor(pokemon$type_1)
#class(pokemon$type_1)
#summary(pokemon$type_1)

pokemon$generation <- as.factor(pokemon$generation)
pokemon$name <- as.factor(pokemon$name)
pokemon$type_2 <- as.factor(pokemon$type_2)
```

How many categorical variables are there? 5 (maybe 6?)

How many continuous variables are there? 7

What are some variables of interest to you?

------------------------------------------------------------------------

## EDA Assignment Prep: \[Insert Data Name Here\]

### Libraries

```{r}
library(here)
library(tidyverse)
library(janitor)
```

### Data

```{r}

```

### Describe the data:

Data author:

Description:

How many observations (rows) are there?

How many variables (columns) are there?

### Summary: Quick overview

```{r}

```

### Categorical vs. Continuous

```{r}

```

How many categorical variables are there?

How many continuous variables are there?

What are some variables of interest to you?
