---
title: "EDA Assignment Prep"
author: "Amber Camp"
format: html
editor: visual
---

## EDA Assignment Prep: Pokemon data

### Libraries

```{r, message = FALSE}
library(here)
library(tidyverse)
library(janitor)
```

### Data

```{r, message = FALSE}
pokemon <- read_csv(here("data/pokemon.csv"))

pokemon <- pokemon %>% 
  clean_names()
```

### Describe the data: Pokemon

Data author: Alberto Barradas

Description: This is pokemon data for 721 pokemon and their stats.

How many observations (rows) are there? 800

How many variables (columns) are there? 13

### skimr: Quick overview

```{r, message = FALSE}
skim_pokemon <- skimr::skim(pokemon)
```

### Summary: Quick overview

```{r}
summary(pokemon)
```

### Categorical vs. Continuous

```{r}

pokemon$number <- as.factor(pokemon$number)
#class(pokemon$number)
#summary(pokemon$number)

pokemon$type_1 <- as.factor(pokemon$type_1)
#class(pokemon$type_1)
#summary(pokemon$type_1)

pokemon$generation <- as.factor(pokemon$generation)
pokemon$name <- as.factor(pokemon$name)
pokemon$type_2 <- as.factor(pokemon$type_2)
```

How many categorical variables are there? 5 (maybe 6?)

How many numeric variables are there? 7

What are some variables of interest to you? how many legendary?

------------------------------------------------------------------------

## EDA Assignment Prep: \[Insert Data Name Here\]

### Data

```{r, message = FALSE}
olympics <- read_csv(here("data/athlete_events.csv"))

olympics <- olympics %>% 
  clean_names()
```

### Describe the data:

Data author: R. Griffin

Description: An extensive culmination of data from the Olympic Games over 120 years.

How many observations (rows) are there? 271,116

How many variables (columns) are there? 15

### Summary: Quick overview

```{r}
summary(olympics)
```

### Categorical vs. Numeric

```{r}
olympics$id <- as.factor(olympics$id)
olympics$year <- as.factor(olympics$year)

class(olympics$id)
class(olympics$year)

skim_olympics <- skimr::skim(olympics)
```

How many categorical variables are there? 12

How many numeric variables are there? 3

What are some variables of interest to you? age, height, and weight

What kinds of questions could you explore with this data?

-   What height/weight/age are the most common among different medal winners?

-   How does an athlete compare to themselves in previous Olympics - in terms of medal placement?
